### 2018-01-30 written_docs/draft.Rmd
Oh goody, start some writing!

---
title             : "Race-based disparities in allocation of discipline are associated with county-level rates of bias"
shorttitle        : "Discipline Disparities"

author: 
  - name          : "Travis Riddle"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : ""
    email         : "triddle@princeton.edu"
  - name          : "Stacey Sinclair"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Princeton University"
  - id            : "2"
    institution   : "Princeton University"

author_note: >

  Enter author note here.

abstract: >
  Enter abstract here.

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
---

```{r include = FALSE}
library("papaja")
```
# Introduction

# Methods
## Data Sources
```{r summary-data}
library(dplyr)
library(forcats)
library(haven)
schools_dat <- read.csv('/Users/travis/Documents/gits/Data/crdc201314csv/CRDC2013_14_SCH.csv')

schools_dat %>%
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, SCH_ENR_HI_M:TOT_ENR_F) %>%
  gather(group, total_number, SCH_ENR_HI_M:TOT_ENR_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>%
  separate(group, into=c('prefix', 'group'), -4) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='NR_', 
                          biracial='TR_', white='WH_')) %>%
  group_by(COMBOKEY, group) %>%
  mutate(total_number = sum(total_number)) %>%
  ungroup() %>%
  select(-prefix, -gender) %>%
  distinct() -> enrollment

enrollment %>% group_by(group) %>% summarise(sum(total_number)) -> enrollment_totals
```
We used three distinct data sources for the work described here. The academic disciplinary data is part of the Civil Rights Data Collection (CRDC) through the US Department of Education. The dataset we used comes from the 2013-2014 academic year and has data on "all public local and educational agencies and schools, including long-term secure juvenile justice facilities, charter schools, alternative schools, and schools serving students with disabilities." In total, the CRDC data represents `r dim(schools_dat)[1]` institutions enrolling approximately 50 million students, of which approximately 25 million are white and 7.8 million are Black.

We obtained county-level demographic information for use as covariates and state-level demographic information for use in post-stratification from the US Census Bureau. For both county-level and state-level demographics, we use 5-year estimatesfor the period ending in 2014 from the American Community Survey, which surveys around 295k households per month.

```{r iat-data}
df <- rbind(read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2013.sav'),
            read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2012.sav'),
            read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2010.sav'))

df2 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2011.sav')
df3 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2009.sav')
df4 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2008.sav')
df5 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2007.sav')
df6 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2006.sav')
df7 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2005.sav')
df8 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2004.sav')
df9 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2002-2003.sav')
df10 <- read_sav('/Users/travis/Documents/gits/Data/iat/Race IAT.public.2014.sav')

"%ni%" <- Negate("%in%")

df %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age)  -> subdat

df2 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all,  
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df3 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df4 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df5 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df6 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df7 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  mutate(raceomb = ethnic) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df8 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  mutate(raceomb = ethnic) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df9 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblacks_0to10) &
           !is.na(age)) %>%
  mutate(raceomb = ethnic) %>%
  mutate(tblack_0to10=tblacks_0to10,
         twhite_0to10=twhites_0to10) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

df10 %>%
  filter(CountyNo!='' &
           !is.na(D_biep.White_Good_all) &
           !is.na(tblack_0to10) &
           !is.na(age)) %>%
  select(CountyNo, STATE, D_biep.White_Good_all, 
         tblack_0to10, twhite_0to10, raceomb, age) %>%
  rbind(subdat) %>%
  filter(raceomb==5 | raceomb==6) -> subdat

subdat %>%
  mutate(explicit_bias=tblack_0to10) %>%
  #filter(!is.na(twhite_0to10)) %>%
  #mutate(explicit_bias = twhite_0to10 - tblack_0to10) %>%
  mutate(age_bin = cut(age, breaks=c(14, 24, 34, 54, 75, 120))) %>%
  mutate(race = fct_recode(as.character(raceomb), 
                           'Black'='5', 'White'='6')) %>%
  filter(!is.na(age_bin)) %>%
  filter(race=='White') %>%
  filter(STATE %ni% 
           c('AA', 'AE', 'AP', 'AS', 'FM', 'GU', 'MH', 'MP', 'PR', 'VI')) %>%
  mutate(county_id = paste(STATE, CountyNo, sep='-')) -> individual_data
```

We used measurements of implicit and explicit bias available from data collected through Project Implicit [@xu2014psychology]. From this data, we used only respondents who had geographic information that would allow us to place them in a United States county, identified as White, and visited the site before 2015. This consisted of approximately 1.1 million total respondents.

## Measures

```{r summary_counts}
schools_dat %>% 
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
         SCH_DISCWODIS_ISS_HI_M:TOT_DISCWODIS_ISS_F, 
         SCH_DISCWDIS_ISS_IDEA_HI_M:TOT_DISCWDIS_ISS_IDEA_F) %>% 
  gather(group, number, SCH_DISCWODIS_ISS_HI_M:TOT_DISCWDIS_ISS_IDEA_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>% 
  separate(group, into=c('prefix', 'group'), -4) %>%
  separate(prefix, into=c('prefix', 'disability'), -3) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='SS_', 
                          total='EA_', biracial='TR_', white='WH_')) -> tempout_susp
%>%
  mutate(disability=fct_recode(disability, not_disabled='_I', disabled='ID',
                               not_disabled='S_', disabled='A_')) %>%
  select(-prefix) -> susp_inschool

susp_inschool %>%
  filter(group=='black'|group=='white') %>%
  group_by(COMBOKEY, LEA_STATE, LEA_NAME, LEAID,
           CCD_LATCOD, CCD_LONCOD, SCH_NAME, group) %>%
  summarise(number = sum(number, na.rm=T)) %>%
  left_join(enrollment) %>%
  mutate(proportion = number/total_number) -> tempout

tempout$number[which(tempout$number<0)] <- NA

tempout %>% 
  ungroup() %>%
  filter(is.finite(proportion)) %>%
  filter(total_number>number) %>%
  mutate(metric='inschool_susp') %>%
  mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
  distinct() -> subdat

###############
schools_dat %>% 
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
         SCH_GRADE_PS, SCH_PSDISC_SINGOOS_HI_M:SCH_PSDISC_SINGOOS_IDEA_F) %>% 
  filter(SCH_GRADE_PS=='YES') %>%
  gather(group, number, SCH_PSDISC_SINGOOS_HI_M:SCH_PSDISC_SINGOOS_IDEA_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>% 
  separate(group, into=c('prefix', 'group'), -4) %>%
  separate(prefix, into=c('prefix', 'disability'), -3) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='OS_', 
                          biracial='TR_', white='WH_')) -> tempout

%>%
  mutate(disability=fct_recode(disability, not_disabled='_I', disabled='ID',
                             not_disabled='GO', disabled='A_')) %>%
  select(-prefix) -> ps_susp

df_school %>% 
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD,
         SCH_GRADE_PS, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>% 
  filter(SCH_GRADE_PS=='YES') %>%
  gather(group, number_2, SCH_PSDISC_MULTOOS_HI_M:TOT_PSDISC_MULTOOS_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>%
  separate(group, into=c('prefix', 'group'), -4) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='OS_', 
                          biracial='TR_', white='WH_')) %>%
  select(COMBOKEY, group, gender, number_2) %>%
  right_join(ps_susp) %>%
  filter(number>=0 & number_2 >= 0) %>%
  mutate(number = number+number_2) -> ps_susp

ps_susp %>%
  left_join(enrollment) %>%
  group_by(COMBOKEY, LEA_STATE, LEA_NAME, 
           CCD_LATCOD, CCD_LONCOD,  SCH_NAME, group) %>%
  summarise(number = sum(number),
            total_number = sum(total_number)) %>%
  mutate(proportion = number/total_number) -> full_dat

full_dat %>% 
  filter(group=='black'|group=='white') %>%
  filter(is.finite(proportion)) %>%
  filter(number>=0) %>%
  filter(total_number>number) %>%
  mutate(group=droplevels(group)) %>%
  ungroup() %>%
  mutate(metric='preschool_susp') %>%
  mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
  rbind(subdat) -> subdat

df_school %>% 
  select(LEA_STATE:LEAID, CCD_LATCOD, CCD_LONCOD, 
         SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>% 
  gather(group, number, SCH_PSDISC_EXP_HI_M:TOT_PSDISC_EXP_F) %>%
  separate(group, into=c('group', 'gender'), -2) %>% 
  separate(group, into=c('prefix', 'group'), -4) %>%
  mutate(group=fct_recode(group, am_indian='AM_', asian='AS_', black='BL_',
                          hispanic='HI_', pac_isl='HP_', total='XP_', 
                          biracial='TR_', white='WH_')) %>%
  select(-prefix) -> ps_exp

ps_exp %>%
  left_join(enrollment) %>%
  group_by(COMBOKEY, LEA_STATE, LEA_NAME, 
           CCD_LATCOD, CCD_LONCOD, SCH_NAME, group) %>%
  summarise(number = sum(number),
            total_number = sum(total_number)) %>%
  mutate(proportion = number/total_number) -> full_dat

full_dat %>% 
  filter(group=='black'|group=='white') %>%
  filter(is.finite(proportion)) %>%
  filter(number>=0) %>%
  filter(total_number>number) %>%
  ungroup() %>%
  mutate(metric='preschool_expulsion') %>%
  mutate(LEA_STATE = droplevels(LEA_STATE)) %>%
  rbind(subdat) -> subdat

```
The primary outcome in this analysis is a count of the number of students by race (black and white) who received one of several types of disciplinary action. Originally, and as specified in our preanalysis plan, our analyses focused on 13 actions - corporal punishment, in-school suspension, out-of-school suspension, expulsion with educational services, expulsion without educational services, expulsion under zero-tolerance policies, referral to law enforcement, school-related arrests, mechanical restraint, physical restraint, seclusion, preschool suspension, and preschool expulsion.  However, upon further study, we discovered reasons we thought justified excluding a number of these outcomes. In particular, seclusion, physical restraint, and mechanical restraint are not disciplinary actions, but are rather used as means to restrain students who are at risk of harming themselves or others. Additionally, the number of preschool students who are expelled or suspended is vanishingly smallseveral of these outcomes are not used for discipline, but rather as forms of treatment for students who pose threats to themselves or others (seclusion, physical restraint, mechanical restraint), and the number of preschool expulsions and suspensions are vanishingly small (10 expulsions, )

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}