### 2018-05-18 written_docs/Riddle_Sinclair_PNAS/pnas_supporting.Rmd

---
title: "PNAS Supporting Information"
author: "Travis Riddle & Stacey Sinclair"
date: "5/11/2018"
output: 
  bookdown::pdf_document2:
    toc: no
    keep_tex: true
bibliography: ref.bib
---

The OSF page for this project (https://osf.io/pu79a/) hosts many additional details. These include interactive maps for all disciplinary outcomes, information for a preregistered version of these analyses, explanations for departures from these plans, code, and aggregated data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(dplyr)
library(knitr)
library(kableExtra)
library(forcats)
library(tidyr)
library(stringr)
library(ggplot2)
```

```{r, reg-coefs, results='asis'}
base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/full_models/'
plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA)

# weighted bias w/exclusions & explicit difference
metrics <- list.files(base_path)
#metrics <- c('expulsion_combined', 'in_school_arrest', 'law_enforcement', 'oos_susp')#1245
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  for (k in files){
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    posterior_combo[,,j] <- t(dat[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }
  data.frame(coef=colnames(dat)[1:30],
             est=rowMeans(p_cons),
             lower=apply(p_cons,1,quantile, probs=.025),
             upper=apply(p_cons,1,quantile, probs=.975),
             p_neg = rowSums(p_cons<0)/4000,
             metric=rep(i, 30)) %>%
    rbind(plot.dat) %>%
    filter(!is.na(est)) -> plot.dat
}

plot.dat$metric <- factor(plot.dat$metric, c('oos_susp', 'inschool_susp',
                                             'law_enforcement', 
                                             'expulsion_combined',
                                             'in_school_arrest'))

plot.dat$metric <- fct_recode(plot.dat$metric, 
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')
plot.dat$coef <- fct_recode(plot.dat$coef,
                            'Intercept' = '(Intercept)',
                            'total population' = 'total_pop',
                            'proportion black' = 'black_prop',
                            'proportion white' = 'white_prop',
                            'black-white ratio' = 'b.w.ratio',
                            'college grads' = 'col_grads',
                            'income' = 'med_income',
                            'poverty' = 'poverty_rate',
                            'unemployment' = 'unemp_rate',
                            'crime' = 'crime_rate',
                            'housing density' = 'density',
                            'mobility' = 'mobility',
                            'dissimilarity' = 'dissim',
                            'total population*race: white' = 'groupwhite:total_pop',
                            'proportion black*race: white' = 'groupwhite:black_prop',
                            'proportion white*race: white' = 'groupwhite:white_prop',
                            'black-white ratio*race: white' = 'groupwhite:b.w.ratio',
                            'college grads*race: white' = 'groupwhite:col_grads',
                            'income*race: white' = 'groupwhite:med_income',
                            'poverty*race: white' = 'groupwhite:poverty_rate',
                            'unemployment*race: white' = 'groupwhite:unemp_rate',
                            'crime*race: white' = 'groupwhite:crime_rate',
                            'housing density*race: white' = 'groupwhite:density',
                            'mobility*race: white' = 'groupwhite:mobility',
                            'dissimilarity*race: white' = 'groupwhite:dissim',
                            'race: white' = 'groupwhite',
                            'implicit bias' = 'weighted_bias',
                            'explicit bias' = 'weighted_explicit_diff',
                            'implicit bias*race: white' = 'groupwhite:weighted_bias',
                            'explicit bias*race: white' = 'groupwhite:weighted_explicit_diff')

plot.dat %>%
  mutate(est = round(est, 2)) %>%
  mutate(estimate=paste(est, ' [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
  mutate(estimate=kableExtra::cell_spec(estimate, 'latex', bold=ifelse(p_neg<.025|p_neg>.975, TRUE, FALSE))) %>%
  dplyr::select(metric, estimate, coef) %>%
  spread(metric, estimate) -> tabledata

tabledata <- tabledata[c(1:7, 23:27, 30, 8, 28, 29, 9:19, 22, 20, 21),]
names(tabledata)[1] <- ''

knitr::kable(tabledata, 
             caption = '\\label{tab:reg-coefs}Regression coefficient estimates for the population-level (i.e. fixed) effects, 
             along with 95\\% uncertainty intervals for each of the disciplinary metrics. Bias estimates are from race data.', 
           format = 'latex', booktabs=T, escape=F, row.names = F) %>%
  landscape() 
```

```{r disc-probs, fig.cap="\\label{fig:disc-probs}Association between bias and the probability of receiving discipline for black and white students. Lines represent the mean of the posterior. Bands indicate 95% uncertainty intervals. Printed p_{pos} represents the posterior probability that the association is positive."}
plot.dat <- data.frame(group=NA, weighted_bias=NA, est=NA, lower=NA, 
                       upper=NA, bias=NA, metric=NA, p_neg_black=NA, p_neg_white=NA)
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  for (k in files){
    #print(i)
    load(paste(pth,k,sep=''))
    df <- as.matrix(m)
    posterior_combo[,,j] <- t(df[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }

  newdat <- expand.grid(group = rownames(table(m$data$group)),
                        total_pop=0,
                        unemp_rate=0,
                        med_income=0,
                        poverty_rate=0,
                        col_grads=0,
                        white_prop=0,
                        black_prop=0,
                        b.w.ratio=0,
                        mobility=0,
                        crime_rate=0,
                        density=0,
                        dissim=0,
                        weighted_bias = seq(-3,3,.25),
                        weighted_explicit_diff = 0,
                        number=0,
                        total_number=1)
  y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)

  cbind(newdat, t(y_hat)) %>% 
    mutate(index=row_number()) %>%
    gather(sample, value, `1`:`4000`) -> posterior_distribution

  p_neg_blackimp <- sum(p_cons[3,]>0)/4000
  p_neg_blackexp <- sum(p_cons[4,]>0)/4000
  p_neg_whiteimp <- sum((p_cons[3,]+p_cons[17,])>0)/4000
  p_neg_whiteexp <- sum((p_cons[4,]+p_cons[18,])>0)/4000

  posterior_distribution %>%
    select(group, weighted_bias, sample, value, index) %>%
    group_by(group, weighted_bias) %>%
    summarise(est = mean(value),
              lower = quantile(value, .025),
              upper = quantile(value, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Implicit',
           metric = i,
           p_neg_black=p_neg_blackimp,
           p_neg_white=p_neg_whiteimp) %>%
    rbind(plot.dat) -> plot.dat

  newdat <- expand.grid(group = rownames(table(m$data$group)),
                        total_pop=0,
                        unemp_rate=0,
                        med_income=0,
                        poverty_rate=0,
                        col_grads=0,
                        white_prop=0,
                        black_prop=0,
                        b.w.ratio=0,
                        mobility=0,
                        crime_rate=0,
                        density=0,
                        dissim=0,
                        weighted_bias = 0,
                        weighted_explicit_diff = seq(-3, 3, .25),
                        number=0,
                        total_number=1)
  y_hat <- rstanarm::posterior_linpred(m, newdata = newdat, re.form=~0, transform=T)

  cbind(newdat, t(y_hat)) %>% 
    mutate(index=row_number()) %>%
    gather(sample, value, `1`:`4000`) -> posterior_distribution

  posterior_distribution %>%
    select(group, weighted_explicit_diff, sample, value, index) %>%
    mutate(weighted_bias = weighted_explicit_diff) %>%
    group_by(group, weighted_bias) %>%
    summarise(est = mean(value),
              lower = quantile(value, .025),
              upper = quantile(value, .975)) %>%
    ungroup() %>%
    mutate(bias = 'Explicit',
           metric=i,
           p_neg_black=p_neg_blackexp,
           p_neg_white=p_neg_whiteexp) %>%
    rbind(plot.dat) -> plot.dat

}

plot.dat$metric <- factor(plot.dat$metric, c('in_school_arrest', 
                                         'expulsion_combined', 
                                         'law_enforcement',
                                         'inschool_susp', 
                                         'oos_susp'))

plot.dat$metric <- fct_recode(plot.dat$metric, 
                               'Out-of\nSchool\nSusp' = 'oos_susp',
                               'In-School\nSusp' = 'inschool_susp',
                               'School\nArrest' = 'in_school_arrest',
                               'Law Enf.\nReferral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')

plot.dat <- plot.dat[!is.na(plot.dat$weighted_bias),]

plot.dat %>%
  filter(weighted_bias==2.5) %>%
  mutate(text_y = est,
         text_x = 2.75) %>%
  select(group, text_y, text_x, metric, p_neg_black, p_neg_white, bias) %>%
  distinct() %>%
  gather(group1, p_neg, p_neg_black, p_neg_white) %>%
  mutate(group1 = str_sub(group1, 7, -1)) %>%
  filter(group==group1) %>%
  select(-group1) -> text_data

text_data$text_y[8] <- text_data$text_y[8]+.00025

text_data$text_comp <- '=='
text_data$text_comp[text_data$p_neg>.99] <- '>'
text_data$text_comp[text_data$p_neg<.01] <- '<'
text_data$p_neg[text_data$p_neg>.99] <- .99
text_data$p_neg[text_data$p_neg<.01] <- .01

plot.dat %>%
  filter(!is.na(weighted_bias)) %>%
  ggplot(aes(x=weighted_bias, y=est, group=group, linetype=group)) +
    geom_line(aes(color=group)) +
    geom_ribbon(aes(ymin=lower, ymax=upper), alpha=.15) +
    theme_classic() +
    geom_text(data=text_data, parse=T, size=2.5, 
              aes(x=2.8, y=text_y, vjust=0,
                  label=paste0('p[pos]', text_comp, round(p_neg,2)))) +
    ylab('Probability of \n disciplinary action') +
    xlab('Standardized Bias') + 
    facet_grid(metric~bias, scales='free') +
    xlim(-2.75, 3.25) +
    theme(axis.title.y = element_text(size=10),
        legend.title = element_blank(),
        legend.position='top',
        axis.text = element_text(size=10),
        axis.title.x = element_text(size=12, face='bold'),
        legend.text = element_text(size=10),
        strip.text.y=element_text(size=7)) -> p1

p1
```

```{r, reg-coefs-sex, results='asis'}
base_path <- '/Users/travis/Documents/gits/educational_disparities/cluster/output/sex_bias/'
plot.dat <- data.frame(coef = NA, est = NA, lower = NA, upper = NA, p_neg = NA, metric = NA)

# weighted bias w/exclusions & explicit difference
metrics <- list.files(base_path)
#metrics <- c('expulsion_combined', 'in_school_arrest', 'law_enforcement', 'oos_susp')#1245
for (i in metrics){
  pth <- paste(base_path, i, '/', sep='')

  files <- list.files(pth)
  j <- 1
  posterior_combo <- array(0, dim=c(30,4000,length(files)))
  for (k in files){
    load(paste(pth,k,sep=''))
    dat <- as.matrix(m)
    posterior_combo[,,j] <- t(dat[,c(1:30)])
    j <- j+1
  }
  p_cons <- parallelMCMCcombine::consensusMCcov(posterior_combo, shuff = T)

  for (param in 1:30){
    m$stanfit@sim$samples[[1]][[param]] <- p_cons[param,1:1000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[2]][[param]] <- p_cons[param,1001:2000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[3]][[param]] <- p_cons[param,2001:3000]
  }
  for (param in 1:30){
    m$stanfit@sim$samples[[4]][[param]] <- p_cons[param,3001:4000]
  }
  data.frame(coef=colnames(dat)[1:30],
             est=rowMeans(p_cons),
             lower=apply(p_cons,1,quantile, probs=.025),
             upper=apply(p_cons,1,quantile, probs=.975),
             p_neg = rowSums(p_cons<0)/4000,
             metric=rep(i, 30)) %>%
    rbind(plot.dat) %>%
    filter(!is.na(est)) -> plot.dat
}

plot.dat$metric <- factor(plot.dat$metric, c('oos_susp', 'inschool_susp',
                                             'law_enforcement', 
                                             'expulsion_combined',
                                             'in_school_arrest'))

plot.dat$metric <- fct_recode(plot.dat$metric,
                               'Out-of-School Suspension' = 'oos_susp',
                               'In-School Suspension' = 'inschool_susp',
                               'School-Related Arrest' = 'in_school_arrest',
                               'Law Enf. Referral' = 'law_enforcement',
                               'Expulsion' = 'expulsion_combined')
plot.dat$coef <- fct_recode(plot.dat$coef,
                            'Intercept' = '(Intercept)',
                            'total population' = 'total_pop',
                            'proportion black' = 'black_prop',
                            'proportion white' = 'white_prop',
                            'black-white ratio' = 'b.w.ratio',
                            'college grads' = 'col_grads',
                            'income' = 'med_income',
                            'poverty' = 'poverty_rate',
                            'unemployment' = 'unemp_rate',
                            'crime' = 'crime_rate',
                            'housing density' = 'density',
                            'mobility' = 'mobility',
                            'dissimilarity' = 'dissim',
                            'total population*race: white' = 'groupwhite:total_pop',
                            'proportion black*race: white' = 'groupwhite:black_prop',
                            'proportion white*race: white' = 'groupwhite:white_prop',
                            'black-white ratio*race: white' = 'groupwhite:b.w.ratio',
                            'college grads*race: white' = 'groupwhite:col_grads',
                            'income*race: white' = 'groupwhite:med_income',
                            'poverty*race: white' = 'groupwhite:poverty_rate',
                            'unemployment*race: white' = 'groupwhite:unemp_rate',
                            'crime*race: white' = 'groupwhite:crime_rate',
                            'housing density*race: white' = 'groupwhite:density',
                            'mobility*race: white' = 'groupwhite:mobility',
                            'dissimilarity*race: white' = 'groupwhite:dissim',
                            'race: white' = 'groupwhite',
                            'implicit bias' = 'weighted_bias_sex',
                            'explicit bias' = 'weighted_explicit_sex',
                            'implicit bias*race: white' = 'groupwhite:weighted_bias_sex',
                            'explicit bias*race: white' = 'groupwhite:weighted_explicit_sex')
plot.dat %>%
  mutate(est = round(est, 2)) %>%
  mutate(estimate=paste(est, ' [', round(lower, 2), ',', round(upper, 2),']', sep='')) %>%
  mutate(estimate=kableExtra::cell_spec(estimate, 'latex', bold=ifelse(p_neg<.025|p_neg>.975, TRUE, FALSE))) %>%
  dplyr::select(metric, estimate, coef) %>%
  spread(metric, estimate) -> tabledata

tabledata <- tabledata[c(1:7, 23:27, 30, 8, 28, 29, 9:19, 22, 20, 21),]
names(tabledata)[1] <- ''

knitr::kable(tabledata, 
             caption = '\\label{tab:reg-coefs-sex}Regression coefficient estimates for the population-level (i.e. fixed) effects, 
             along with 95\\% uncertainty intervals for each of the disciplinary metrics. Bias estimates are from sexuality data.', 
           format = 'latex', booktabs=T, escape=F, row.names = F) %>%
  landscape()

```

```{r cor-tab}
county_means <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means.csv')
county_means_sex <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/county_means_sexuality.csv')
covs <- read.csv('/Users/travis/Documents/gits/educational_disparities/output/covariates.csv')
county_means %>%
  mutate(race_implicit = weighted_bias,
         race_explicit = weighted_explicit_diff) %>%
  select(county_id, race_implicit, race_explicit) %>%
  left_join(county_means_sex) %>%
  mutate(sex_implicit = weighted_bias,
         sex_explicit = weighted_explicit) %>%
  select(county_id, race_implicit, race_explicit, sex_implicit, sex_explicit) %>%
  full_join(covs) %>%
  filter(county_id!='NA-NA') -> preds

preds %>% 
  select(race_implicit:sex_explicit, total_pop:poverty_rate, crime_rate:dissim) %>%
  cor(., use='pairwise.complete.obs') -> cormat

preds %>% 
  mutate(crime_rate = crime_rate*1000, 
         total_pop = total_pop/1000,
         white_pop = white_pop/1000,
         black_pop = black_pop/1000,
         med_income = med_income/1000) %>%
  select(race_implicit:sex_explicit, total_pop:poverty_rate, crime_rate:dissim) %>%
  summarise_all(funs(mean(., na.rm=T))) -> means

preds %>% mutate(crime_rate = crime_rate*1000,
                 total_pop = total_pop/1000,
                 white_pop = white_pop/1000,
                 black_pop = black_pop/1000,
                 med_income = med_income/1000) %>%
  select(race_implicit:sex_explicit, total_pop:poverty_rate, crime_rate:dissim) %>%
  summarise_all(funs(sd(., na.rm=T))) -> sdevs

upper<-round(cormat, 2)
upper[upper.tri(cormat, diag = T)]<-""
upper<-as.data.frame(upper)
upper %>%
  mutate_all(funs(as.character)) %>%
  rbind(round(means, 2)) %>%
  rbind(round(sdevs, 2)) -> cortab

vars <- c("1) Implicit Race", "2) Explicit Race", "3) Implicit Sex", 
          "4) Explicit Sex", "5) Population Size", "6) White Population",
          "7) Black Population", "8) College Grads", "9) Unemployment", 
          "10) Income", "11) Poverty", "12) Crime", "13) Housing Density", 
          "14) Mobility", "15) Proportion White", "16) Proportion Black", 
          "17) Black-White Ratio", "18) Segregation", "Mean", "SD")

cortab <- cbind(vars, cortab)

kable(cortab, 
      caption = '\\label{tab:cor-tab}Correlation matrix for all county-level predictor variables.',
      format = 'latex', booktabs=T, escape=F, row.names = F,
             col.names=c('', seq(1:18))) %>%
  landscape() %>%
  kable_styling(latex_options=c('scale_down')) %>%
  footnote(number = c("Crime rate is scaled to rate per 1000 people",
                      "Total population, black population, and white population are in 1000's of people",
                      "Income is in 1000's of dollars")) %>%
  group_rows("", 19, 20, latex_gap_space = ".5em")
```